---
title: "The State of Knitting"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(dplyr)
library(knitr)
library(readr)
library(mltools)
library(naniar) #for replace_with_na
library(naivebayes)
library(class)
library(caret)
library(MLmetrics)
library(nnet)
library(SmartEDA) #quick views
library(lubridate) #grouping by start month
library(corrplot)
library(ggcorrplot)
library(maps)
```

## Sidebar {.sidebar}

Knitting is an incredibly popular hobby. Websites such as Ravelry.com allow pattern designers to sell their patterns directly to knitters. This dashboard captures different aspects of pattern and pattern author data scraped from ravelry.com

Preparing the dashboard based on the instructions given at @magallanes_reyes_data_2022, and @aden-buie_flexdashboard_2023. 

You can find the code for this dashboard here: https://github.com/elizageeslin/datavisualization

## Column {.tabset .tabset-fade}

### Univariate Categorical Plot

```{r}
#reading in the data for univariate plot

plot1data <- read_csv("item_type_bar_graph.csv", show_col_types = FALSE)
```


```{r}
#plot 1
titleText='What are the most popular pattern types?'
subtitleText='Count of the most popular patterng grouped by the type of item the pattern makes'
sourceText='Source: Ravelry.com'
x.AxisText="Item Type"
y.AxisText="Count of Patterns"

ggplot(data = plot1data, aes(x = reorder(item_type, desc(count)), y = count)) +
        geom_bar(fill ="#2E475D", stat = 'identity') + 
        labs(title=titleText
             , subtitle = subtitleText
             , x = x.AxisText
             , y = y.AxisText
             , caption = sourceText
             , color="#213343") +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)
              , panel.background = element_rect(fill = "#F6F9FC")) +
        geom_text(aes(label = count)
                    , vjust = 1.5
                    , size = 3
                    , color="#F6F9FC") +
        geom_vline(xintercept = 4.5, color="#FF7A59", linetype = "dashed", size = 1) + 
        geom_text(aes(label = 'Top 4 Categories Account for more than 50% of patterns.')
            , x = 4.5
            , y = 280
            , vjust = -0.5
            , hjust = -0.05
            , size = 2.5
            , color = "#FF7A59") 
```

### Numerical Histogram Plot

```{r}
#reading in the data for plot3

plot3data <- read_csv("price_histogram.csv", show_col_types = FALSE)

# plot3 - filter currency to usd only and filter out all free patterns

plot3data <- plot3data %>% filter(currency == "USD" & !is.na(price))

# Calculate the mean price from the data
mean_price <- mean(plot3data$price)
```

```{r}
titleText='What is the distribution of pattern price?'
subtitleText='For patterns that cost money, this graph looks at the distribution of prices in USD'
sourceText='Source: Ravelry.com'
x.AxisText='Pattern Price ($)'
y.AxisText='Density'

ggplot(data = plot3data, aes(x=price)) + 
    geom_histogram(aes(y = after_stat(density)), bins = 15, fill = "#516F90") +
    labs(title=titleText
            , subtitle = subtitleText
            , x = x.AxisText
            , y = y.AxisText
            , caption = sourceText
            , color="#213343") + 
    theme(panel.background = element_rect(fill = "#F6F9FC")) +
    geom_vline(xintercept= mean(plot3data$price), color="#FF7A59", linetype = "dashed", size = 1) + 
    geom_text(aes(label = 'Mean Price = $7.12 USD')
            , x = 7
            , y = 0.225
            , vjust = -0.5
            , hjust = -0.09
            , size = 3.0
            , color = "#FF7A59") 
```

### Categorical x Categorical Plot

```{r}
#reading in the data for cat-cat plot

plot2data <- read_csv("item_type_free_bar_graph.csv", show_col_types = FALSE)
```

```{r}
#plot 2
titleText='Breakdown of Free vs. Paid Patterns by Type'
subtitleText='Not all pattern costs money; For the top 10 item types, what is the split between free patterns and paid patterns?'
x.AxisText='Item Type'
y.AxisText='Free or Paid (%)'
fillText = 'Pattern Cost'
sourceText='Source: Ravelry.com'


ggplot(data = plot2data, aes(fill = cost, x = reorder(item_type, desc(count)), y = count)) +
        geom_bar(position="fill", stat="identity") + 
        labs(title=titleText
             , subtitle = subtitleText
             , caption = sourceText
             , x = x.AxisText
             , y = y.AxisText
             , fill = fillText
             , color="#213343") +
        scale_fill_manual(values = c("#0FBFBF", "#FF7A59")) + 
        theme(axis.text.x = element_text(angle = 45, hjust = 1)
              , panel.background = element_rect(fill = "#F6F9FC")) +
         geom_text(aes(label = percent_label)
                   , position = position_fill(vjust = 0.5)
                   , color="#213343")
```


### Map Plot

```{r}
#reading in the data for map

map_data <- read_csv("map_data.csv", show_col_types = FALSE)
world_coordinates <- map_data("world") 

map_data <- merge(world_coordinates, map_data, by="region", all = TRUE)

# Sort by group, then order
map_data <- arrange(map_data, group, order)
```

```{r}
# Calculate the center of each state for label placement
country_centers <- map_data %>%
  drop_na(label_lat) %>%
  select('code','label_lat', 'label_long', 'authors_per_capita')

# Add a dummy 'group' column to state_centers
country_centers$group <- 1

country_centers <- country_centers %>%
  distinct()
```


```{r,fig.height=6,fig.width=10}
titleText='Authors of Top Patterns per Capita'
subtitleText='A look at the number of pattern authors per capita by country'
x.AxisText="Latitude"
y.AxisText="Longitude"
sourceText='Nordic countries (Denmark, Finland and Norway) have the highest number of authers per capita. (Source: Ravelry.com)'
fillText ='Authors per capita'

# Create the choropleth map
ggplot(map_data, aes(x = long, y = lat, group = group, fill = authors_per_capita)) +
  geom_polygon(color = "#99AFC4") +
  scale_fill_gradient(low = "#FCEAD7", high = "#FF7A59", na.value = "#DBE4ED") +
  labs(title = titleText
       , subtitle = subtitleText
       , caption = (sourceText)
       , x = x.AxisText
       , y = y.AxisText
       , fill = fillText
       , color="#213343") +
  theme(legend.position = c(0.05, 0.05) 
      , legend.justification = c("left", "bottom")
      , legend.box.just = "left"
      , legend.margin = margin(6, 6, 6, 6)
      , panel.background = element_rect(fill = "#F6F9FC")) +
  geom_text(data = country_centers
      , aes(x = label_long, y = label_lat, label = code)
      , size = 1
      , color = "#213343")

```